---
title: "Challenge 9"
author: "Kaleigh Chi"
editor: visual
format: 
  html: 
    self-contained: true
    code-tools: true
execute: 
  echo: true
  error: true
  warning: false
  messages: false
---

# Challenge 9: Baby Names

```{r}
library(tidyverse)
library(here)
library(gt)
library(DT)

namesA <- read_csv(here("Week 9", "Lab 9", "StateNames_A.csv"))
```

# 3: Summarizing & Visualizing the Number of Allisons

## 3.1 Summarizing the Number of Allisons

```{r}
namesA %>%
  filter(Name == 'Allison') %>%
  rename(Sex = Gender) %>%
  group_by(Sex, State) %>%
  summarize(count_eval = sum(Count),
            .groups = "drop") %>%
  pivot_wider(names_from = Sex,
              values_from = count_eval,
              values_fill = 0) %>%
  gt() %>%
  as.data.frame() %>%
  datatable()
```

## 3.2 Allison Data Set

```{r}
allison_f <- namesA %>%
  filter(Name == 'Allison', Gender == 'F') %>%
  rename(Sex = Gender) %>%
  group_by(Sex, Year) %>%
  summarize(count_eval = sum(Count),
            .groups = 'drop') %>%
  pivot_wider(names_from = Sex,
              values_from = count_eval,
              values_fill = 0)
```

## 3.3 Popularity of Allison Visualization

```{r}
allison_f %>%
  mutate(
    Year = factor(Year)
  ) %>%
  
ggplot(mapping = aes(x = Year, y = F)) +
    geom_bar(stat = "identity") +
    labs(y = NULL,
         x = "Year",
         title = "Frequency of U.S. Babies Named 'Allison' from 1997 to 2014") +
    scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
    theme(plot.title.position = "plot") 
```

# 4: Modeling the Number of Allisons

## 4.4 Modeling the Number of Allisons

```{r}
ggplot(data = allison_f,
        mapping = aes(x = Year, y = F)) +
   geom_point() +
   stat_smooth(method = "lm", formula = y ~ x) +
   labs(y = NULL,
        x = "Year",
        title = "Linear Model of U.S. Babies Named 'Allison' from 1997 to 2014") +
   theme(plot.title.position = "plot")
```

## 4.5 Estimated regression equation

```{r}
estReg <- allison_f |>
  lm(F ~ Year, data = _)

broom::tidy(estReg) %>%
  gt()
```

predictedAllisonsBorn = 209689.7609 - 101.5191 \* year

## 4.6 Plot Residuals of the Model

```{r}
modResid <- allison_f |>
  lm(F ~ Year, data = _)

modResid |>
  broom::augment() |>
  ggplot(mapping = aes(x = .fitted, y = .resid)) +
  geom_point() +
  labs(x = "fitted", y = "resid")
```

Our residual plot appears to be in a nonlinear U-shaped pattern. The residuals tend to be within 500 of the actual values, but I can't say there's equality of variance. This indicates to me that the regression model would not be appropriate.

## 4.7 Allison Model Conclusion

I see that the linear regression model there is a downward slope from 1997 to 2014. Given the nonlinear residual patterns from the residual model, I conclude from this model that there is low confidence that Allison is no longer a popular name.

# 5. Spelling by State

## 5.8 Narrow A Name Data Set

```{r}
aNames_df <- namesA %>%
  filter(Name %in% c("Allan", "Alan", "Allen"), 
         Gender == "M") %>%
  group_by(Name, Year) %>%
  summarise(Count = sum(Count), .groups = 'keep') %>%
  pivot_wider(names_from = Name,
              values_from = Count)
```

```{r}
ggplot(data = aNames_df,
       mapping = aes(x = Year)) +
    geom_line(aes(y = Alan, color = "green")) +
    geom_line(aes(y = Allan, color = "red")) +
    geom_line(aes(y = Allen, color = "blue")) +
    labs(y = NULL,
         x = "Year",
         title = "Popularity of Male 'A' Names From 1997 to 2014") +
    theme(plot.title.position = "plot",
          legend.position = "none") +
    annotate("text", y = 2350, x = 2014, label = "Alan", color = "green") +
    annotate("text", y = 1000, x = 2014, label = "Allen", color = "red") +
    annotate("text", y = 450, x = 2014, label = "Allan", color = "blue")
```

## 5.9 Tables of Baby "Allan" Born in 2000

```{r}
# learned how to use tab_header() with source:
# https://gt.rstudio.com/reference/tab_header.html

# learned how to use cols_label() with source:
# https://gt.rstudio.com/reference/cols_label.html 

namesA %>%
  filter(Name %in% c("Allan", "Allen", "Alan"),
         State %in% c("CA", "PA"), 
         Year == 2000, 
         Gender == "M") %>%
  group_by(Name, State) %>%
  summarise(Count = sum(Count), .groups = 'keep') %>%
  pivot_wider(names_from = Name,
              values_from = Count,
              values_fill = 0) %>%
  gt() %>%
  cols_label(State = "State") %>%
  tab_header(title = "Frequency of Baby 'Alan' Variations",
             subtitle = "California and Pennsyvlania in Year 2000")
```

## 5.10 Convert Counts to Percents

```{r}
# learned how to use tab_options() with sources: 
# https://gt.rstudio.com/reference/tab_options.html
# https://www.liamdbailey.com/post/making-beautiful-tables-with-gt/

namesA %>%
  filter(Name %in% c("Allan", "Allen", "Alan"),
         State %in% c("CA", "PA"),
         Year == 2000,
         Gender == "M") %>%
  group_by(State) %>%
  mutate(Prop = Count / sum(Count)) %>%
  select(-Count, -Year, -Gender) %>%
  pivot_wider(names_from = Name,
              values_from = Prop,
              values_fill = 0) %>%
  gt(groupname_col = "State") %>%
  fmt_percent(columns = 2:4, decimals = 2) %>%
  tab_header(title = "Percentage of Baby 'Alan' Variations Within States",
             subtitle = "California and Pennsylvania in Year 2000") %>%
  tab_options(heading.title.font.size = 20,
              table.background.color = "steelblue",
              heading.align = "left")
```

# Reflections

I looked over my challenge 9 since I noticed I received an incomplete. I made the same corrections to challenge 9 that I did to my lab 9. I changed the Question 5 to make sure the predictedAllisonsBorn was denoted as the y. This is to accurately represent the y for the regression line and it's good practice to make that note in my equation.

I also changed my Question 10 table because I noticed the percentages weren't adding correctly within each states. Initially, my table was finding proportions of each name based on the total table. I realized I left out a group_by(). I made sure to group by states to ensure the accurate proportions were found within each state and thus accurate percentages.
