---
title: "PA 4"
format: html
editor: visual
execute: 
  warning: false
---

# Data Import

## Problem 1: Modify code to read military expenditures data into work space

```{r}
library(readxl) 
library(tidyverse)


military <- read_xlsx(here::here("Week 4", 
                                 "Practice Activity", 
                                 "gov_spending_per_capita.xlsx"), 
                      sheet = "Share of Govt. spending", 
                      skip = 7, 
                      n_max = 190)
```

# Data Cleaning

## Problem 2: Find two methods and write code to replace values with NA

```{r}

military_clean <- military |>
 mutate(across(`1988`:`2019`, .fns = na_if, ". ."))

  
```

## Problem 3: Mutate these columns to all be numeric data types

```{r}
military_clean <- military_clean |>
mutate(across(`1988`:`2019`, .fns = as.numeric))

```

## Incorrect Regions

```{r}
cont_region <- read_xlsx(here::here("Week 4", 
                                    "Practice Activity", 
                                    "gov_spending_per_capita.xlsx"), 
                      sheet = "Regional totals", 
                      skip = 14) |> 
  filter(Region != "World total (including Iraq)", 
         Region != "World total (excluding Iraq)")
```

## Problem 4: Use the anti_join() function to filter out the Country values we don't want in the military_clean data set

```{r}
military_clean <- military_clean |> 
  anti_join(cont_region, by = c("Country" = "Region"))
```

# Data Organization

## Problem 4: Pivot the cleaned up military data set to a longer orientation

```{r}
military_long <- military_clean |>
  pivot_longer(cols = '1988':'2019', names_to = "Year", values_to = "Spending")

```

# Data Visualization Exploration

## Problem 5: Create side-by-side boxplots of military spending for each year

```{r}
ggplot(data = military_long, 
       aes(x = Spending, y = Year, group = Year)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(color = "tomato", alpha = 0.2)
```
